---
title: How to make this website
author: Sean D McAtee
date: '2021-02-21'
slug: []
categories: []
tags: []
Description: ''
Tags: []
Categories: []
DisableComments: no
---

### Motivation

### Other Similar Options & Variations
https://about.gitlab.com/stages-devops-lifecycle/continuous-integration/
https://pages.github.com/
https://www.freecodecamp.org/news/hugo-vs-jekyll-battle-of-static-site-generator-themes/



### Sources:
https://bookdown.org/yihui/blogdown/
https://gohugo.io/


## Overview
https://bookdown.org/yihui/blogdown/workflow.html # how to setup a website with more clicking and less typing

Static website: All HTML files are fully rendered locally and served without any additional changes. This can make hosting & serving simple websites much cheaper and faster. Sites include CSS and JS, so they can be 

Hugo: A static website generator with no original connection to the R universe. You can use Hugo directly with the command line if you don't need to include any code chunks.

Blogdown:



### Installation
https://bookdown.org/yihui/blogdown/installation.html

```{r eval=FALSE}
install.packages("blogdown")
library(blogdown)
blogdown::install_hugo()
```

### Themes
https://bookdown.org/yihui/blogdown/other-themes.html
https://themes.gohugo.io/

Hugo generates sites using 'themes' aka rendering templates. There are many themes available on Hugo, however some of them must be modified to be rendered by `rmarkdown`. There is plenty of documentation if you find a theme that must be modified to work.

After choosing a theme you don't need to install it by hand, it is better if you pass the theme name to `blogdown::new_site()` or by building a new site in an RStudio project context (shown below).


## New Project
https://support.rstudio.com/hc/en-us/articles/200526207-Using-Projects
https://bookdown.org/yihui/blogdown/rstudio-ide.html

You don't have to start a new project in R Studio to build a site, but it keeps everything much more organized if you do.

`File -> New Project -> New Directory` then select "Website using blogdown"
This is where you can name you new directory and set your Hugo theme.


The blogdown author recommends enabling the option "Preview site after building" and disabling the option "Re-knit current preview when supporting files change". To find those options look on R Studio under `Tools -> Project Options...-> Build Tools`


# What just happened?
* You just made an R Studio project


# config.toml
https://bookdown.org/yihui/blogdown/configuration.html#toml-syntax

Website-wide configuration can be done by modifying config.toml

If you are doing anything fancy you will end up in config.toml. It is located in the home directory of your project.

```
## Social links
# use 'fab' when brand icons, use 'fas' when standard solid icons.
[[params.socialIcons]]
icon = "fab fa-linkedin"
title = "Linkedin"
url = "https://www.linkedin.com/in/sean-david-mcatee/"

[[params.socialIcons]]
icon = "fab fa-github"
title = "GitHub"
url = "https://github.com/smcatee/"

[[params.socialIcons]]
icon = "fas fa-envelope"
title = "e-mail"
url = "mailto:sean.d.mcatee@gmail.com"

## Menu items

[menu]

  [[menu.main]]
  name = "Portfolio"
  identifier = "home"
  weight = 100
  url = "/"

  #[[menu.main]]
  #name = "Posts"
  #weight = 200
  #identifier = "posts"
  #url = "/post/"

  [[menu.main]]
  name = "About Me"
  weight = 300
  identifier = "about"
  url = "/about/"

  #[[menu.main]]
  #name = "Contact"
  #weight = 400
  #identifier = "contact"
  #url = "/contact/"
```

# Publishing for the first time

Restart R

```{r eval=FALSE}
blogdown::hugo_build() # This will build the site as if from scratch, all webpages will be rendered
```

After making new pages or any changes

```{r eval=FALSE}
blogdown::serve_site() # This will only render the webpages that have been changed
```


# Get netlify and GitHub setup
https://github.com/
https://www.netlify.com/ # can use GitHub account to sign up for Netlify

On GitHub create a new repository for you site. Make sure that it is Public.
```{bash eval=FALSE, echo=FALSE}
# Some reminders, make sure you know what these git commands do.
git init
git add .
git status
git commit -m "first commit"
git branch -M main
git remote add origin https://github.com/your-username/repository-name.git
git push -u origin main
```


Now that your website is hosted on GitHub, go on Netlify `New Site from Git -> Continuous Deployment [GitHub] -> Pick a Repository -> ...`
The default domain name is <name.netlify.app> but there are options to purchase or set up your a domain name.

# Update config.toml with domain name
Use the domain name generated by Netlify, or use the custom domain name that you setup in Netlify.

In config.toml update the parameter at the top `baseurl = "https://my-website.netlify.app/"`

Don't forget to commit and push your updated website to GitHub.
```{bash eval=FALSE, echo=TRUE}
git add .
git status
git commit -m "updated baseurl"
git push -u origin main
```

Now wait a few seconds, check your Netlify website, fix bugs, edit CSS, make [new posts](/2021/02/21/how-to-make-this-webpage./)
