---
title: How to make this webpage
author: Sean D McAtee
date: '2021-02-21'
slug: []
categories: []
tags: []
Description: ''
Tags: []
Categories: []
DisableComments: no
---

</br>

### Motivation for Rmd.

https://rmarkdown.rstudio.com/gallery.html


The primary characteristics that make Rmd. convenient:

* Simplicity

* Code Blocks

</br></br>

### Libraries Used
```{r}
library(blogdown); library(magrittr)
```

</br></br>


### Open the website project directory in RStudio

From RStudio: `File -> Recent Projects`

The R project will save environment variables from your last session, so everything will load up and ready to go.


</br></br>

### Make a _new_ new webpage

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy('r')
```

```{r eval=FALSE, class.source='klippy'}
blogdown::new_post(title = "How to make this webpage", author = "Sean", ext = ".Rmd")
```

</br>

Add content...

[Rmd cheatsheet](https://pkgs.rstudio.com/rmarkdown/articles/rmarkdown.html#cheatsheets-1)


</br></br>


### Code blocks

The HTML and CSS documents are generated from the Rmd. template by the `blogdown` package which depends upon the `knitr` package. `knitr` can execute and display blocks of code nested in the Rmd. document. R scripts can be executed in Rmd. by `knitr`, as well as [other languages](https://yihui.org/knitr/demo/engines/)). 

```{r class.source='klippy'}
knitr::knit_engines$get() %>% names()
```

</br>


These languages are available for [syntax highlighting and execution in blogdown](https://bookdown.org/yihui/rmarkdown/language-engines.html).

For block execution there are language specific dependencies that work with `knitr`.

</br>

This is what code blocks will look like in Rmd.

```
\```{r}  
#R Block  
\```  
```

</br>

```
\```{python}  
#Python Block  
\```  
```

</br>

```
\```{julia}  
#Julia Block  
\```  
```

</br>

```
\```  
#Default Block  
\```  
```

</br>

Inline code will look like this \``r` `<code>`\` and can be executed by R. If you wish to only highlight without executing then make an inline block without 'r'.

e.g. \` `1 + 2` \` ->becomes-> `1 + 2`

e.g. \``r` `1 + 2`\` ->becomes-> `r 1 + 2`

e.g. \``r` `c(1,2,3) %>% mean()`\` ->becomes-> `r c(1,2,3) %>% mean()`

e.g. \``r` `colnames(PlantGrowth)`\` ->becomes-> `r colnames(PlantGrowth)`

e.g. \``r` `PlantGrowth$weight[PlantGrowth$group == "trt1"] %>% mean()`\` ->becomes-> `r PlantGrowth$weight[PlantGrowth$group == "trt1"] %>% mean()`


</br></br>

### Code block options

Document-wide code block options are defined at the top of Rmd. documents.

</br>

Here is an example from this Rmd. document:

```
---
title: How to make this webpage
author: Sean D McAtee
date: '2021-02-21'
slug: []
categories: []
tags: []
Description: ''
Tags: []
Categories: []
DisableComments: no
---
```

Note: Normally Rmd. documents will contain an `output:` option. This is excluded when using `blogdown`.

</br>

Block options can be defined for specific blocks within the braces (`{}`) at the beginning of the block.

e.g. `{r echo=TRUE, eval=FALSE, hold=TRUE, fig.asp=0.3}`

</br></br>

### HTML, CSS, and JS

The file that `blogdown` outputs is a single `index.html` file in the directory in the name of the webpage. There will also be modifications to some of the files in the site-wide directories.

</br>

It is possible to preserve HTML elements through Rmd. documents.

</br> </br> </br>

The spacing above was made with `</br> </br> </br>` written directly in the text of this .Rmd document.


### JS

JavaScript can be executed in a code block.

```{js, eval=TRUE, echo=TRUE}
$('.body').css('color', 'red')
```

### CSS

CSS can also be included in code blocks.

```{css, eval=FALSE, echo=TRUE}
body {
  color: red;
}
```

</br></br>

### Add content & render changes

```{r eval=FALSE, class.source='klippy'}
blogdown::serve_site()
```

</br></br>

### Commit and push to GitHub!

Push your changes to the main repository branch (the one on GitHub) and see your changes.

</br>

```{bash eval=FALSE, echo=TRUE}
git add .
git status
git commit -m "updated baseurl"
git push -u origin main
```

