---
title: How to make this webpage.
author: Sean D McAtee
date: '2021-02-21'
slug: []
categories: []
tags: []
Description: ''
Tags: []
Categories: []
DisableComments: no
---

## Libraries
```{r}
library(blogdown); library(magrittr)
```

Open Project if not already open.

## Set Options (optional)
```{r eval=FALSE}
options(blogdown.ext = ".Rmd", blogdown.author = "Sean D McAtee")
```


## To make a new webpage
```{r eval=FALSE}
blogdown::new_post(title = "How to make this webpage.")
```

</br> </br>

### Save == Render

</br> </br>

## Rendering the webpage and serving site
```{r eval=FALSE}
blogdown::serve_site()
```

</br></br>


## Code chunks

\```{r}  
#R Chunk  
\```  

\```{python}
#Python Chunk
\```

\```{julia}
#Julia Chunk
\```


```{r}
knitr::knit_engines$get() %>% names()
```

These languages are available for [syntax highlighting and execution](https://bookdown.org/yihui/rmarkdown/language-engines.html).

## HTML

</br> </br> </br>

The spacing above was made with `</br> </br> </br>` passed directly in the .Rmd document.

## JS

```{js, eval=FALSE, echo=TRUE}
$('.title').css('color', 'red')
```

## CSS

```{css, eval=FALSE, echo=TRUE}
body {
  color: red;
}
```


## Making it real
```{r eval=FALSE}
blogdown::serve_site()
```

## Commit and push to GitHub

```{bash eval=FALSE, echo=TRUE}
git add .
git status
git commit -m "updated baseurl"
git push -u origin main