---
title: How to make this webpage
author: Sean D McAtee
date: '2021-02-21'
slug: []
categories: []
tags: []
Description: ''
Tags: []
Categories: []
DisableComments: no
---

</br>

### Motivation for Static Websites

The primary characteristics that make static websites convenient.

* Simple Structure


</br></br>

### Resources

https://pkgs.rstudio.com/rmarkdown/articles/rmarkdown.html#cheatsheets-1

</br>

### Libraries
```{r}
library(blogdown); library(magrittr)
```

</br></br>


### To make a new webpage
```{r eval=FALSE}
blogdown::new_post(title = "How to make this webpage.")
```


</br> </br>

### Rendering the webpage and serving site
```{r eval=FALSE}
blogdown::serve_site()
```



</br></br>


### Code chunks

The HTML and CSS documents are generated from the Rmd. template by the `blogdown` package which depends upon the `knitr` package. `knitr` can execute and display blocks of code nested in the Rmd. document. R scripts can be executed in Rmd. by `knitr`, as well as other languages. ([incl. Python, Ruby, ...](https://yihui.org/knitr/demo/engines/))


Blocks of code can be displayed in highlighted syntax or shaded blocks depending upon the Hugo theme.


```
\```{r}  
#R Chunk  
\```  
```

```
\```{python}  
#Python Chunk  
\```  
```

```
\```{julia}  
#Julia Chunk  
\```  
```


```{r}
knitr::knit_engines$get() %>% names()
```

These languages are available for [syntax highlighting and execution](https://bookdown.org/yihui/rmarkdown/language-engines.html).

For chunk execution there are language specific dependencies that work with `knitr`.

### HTML, CSS, and JS

The file that `blogdown` outputs is a single `index.html` file in the directory in the name of the webpage. There will also be modifications to some of the files in the site-wide directories.

</br>

It is possible to preserve HTML elements through Rmd. documents.

</br> </br> </br>

The spacing above was made with `</br> </br> </br>` written directly in the text of this .Rmd document.


### JS

JavaScript can be executed in a code chunk.

```{js, eval=TRUE, echo=TRUE}
$('.title').css('color', 'red')
```

### CSS

CSS can also be included in code chunks.

```{css, eval=FALSE, echo=TRUE}
body {
  color: red;
}
```


### Making it real

`blogdown` must now generate menu items and connect other dots. The command `serve_site` will prepare the website and serve it locally. This command does the minimal amount of updating to generate the website, unlike `build_site` which will generate the whole site from scratch.

```{r eval=FALSE}
blogdown::serve_site()
```

If the page looks good then we can put it online.


### Commit and push to GitHub!

The website will be hosted on GitHub in a repository. We can now initiate a new git repository in this R project directory.

```{bash eval=FALSE, echo=TRUE}
git add .
git status
git commit -m "updated baseurl"
git push -u origin main
```

